# Miscellanious plot generation for manuscript

NOTE: This began as a .R document before converting to .Rmd for greater readability and easier code segmentation across chunks


```{r SETUP chunk}

#Declare plot-friendly names for variables included in analyses
recodes=c("Time_to_death_or_last_followup_years"="Disease duration",
"Age_at_onset_years"="Age of onset",
"CTYDEL"="Diagnostic delay",
"Sex_at_birth" = "Sex",
"Site_of_Onset" = "Site of onset",
"Phenotype_1" = "Clinical diagnosis (ALS)",
"Phenotype_2" = "Clinical diagnosis (PLS)",
"Phenotype_3" = "Clinical diagnosis (PMA)",
"c9_status" = "C9orf72 expansion",
"ALScc_SBayesR" = "ALS PRS",
"FTDcc_SBayesR" = "FTD PRS",
"SZcc_SBayesR" = "Schizophrenia PRS",
"PDcc_SBayesR" = "Parkinson's disease PRS",
"AZcc_SBayesR" = "Alzheimer's disease PRS",
"AutoProt" = "Autophagy & proteostasis",
"CytoTransp" = "Cyt. Dynam. & Ax. Transp.",
"RNAFunc" = "RNA function")

```

### Generate upset plots for missingness

```{r Missingness plots}
library(dplyr)
library(data.table)
library(ggplot2)
library(optparse)
library(naniar)
data <- fread("/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/JNT_JanFinal_savedata_C5/completesample/saved_mplusfit.tsv")

#Flag the originating cohort for each sample - in the 'new_num_id' column, all project mine samples begin with 999 and all strength samples begin with 555.
data <- data %>%
  mutate(cohort=case_when(grepl("^999",new_num_id) ~ "Project MinE",
                          grepl("^555",new_num_id) ~ "STRENGTH",
                          TRUE ~ NA_character_
                          ))

#Extract N's per country and by 
#table(data$origin,data$cohort)

### Extract missingness profiles
miss_data<- data %>%
  dplyr::select(cohort,Time_to_death_or_last_followup_years,Age_at_onset_years,Diagnostic_delay_years,Sex_at_birth,Site_of_Onset,Phenotype) %>%
  rename(`Disease Duration`=Time_to_death_or_last_followup_years,
         `Age of onset`=Age_at_onset_years,
         `Diagnostic delay`=Diagnostic_delay_years,
         `Sex`=Sex_at_birth,
         `Site of onset`=Site_of_Onset,
         `Clinical diagnosis`=Phenotype)

PM_missplot<- miss_data %>%
  dplyr::filter(cohort=="Project MinE") %>%
  naniar::gg_miss_upset(.)

STR_missplot<- miss_data %>%
  dplyr::filter(cohort=="STRENGTH") %>%
  naniar::gg_miss_upset(.)

#Save in 100 x 160mm files
pdf("~/Downloads/upsetplotPM.pdf",height=3.94,width=6.3)
PM_missplot
dev.off()

pdf("~/Downloads/upsetplotSTR.pdf",height=3.94,width=6.3)
STR_missplot
dev.off()

```


### Generate the 3-panel roc plot

```{r}

singleROCS <- list(clinicalOnlyimp=readRDS("/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/Figures/sourcedata/ROC_curves/OnePanel_roc_Pheno.Rds"),
                   withrarendPRS=readRDS("/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/Figures/sourcedata/ROC_curves/OnePanel_roc_Genetic.Rds"),
                   restrictedClinical=readRDS("/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/Figures/sourcedata/ROC_curves/OnePanel_roc_filtered.Rds"))

singleROCS <- lapply(singleROCS,function(x){x+
    guides(color=guide_legend(ncol=1, byrow=TRUE))+
    theme(legend.position = c(0.775,0.3),# "bottom",
           legend.background = element_rect(fill = "white", color = "black"),
           legend.text = element_text(size=8),
           legend.title = element_text(size=10),
         legend.key.height = unit(0.2, 'cm'),
           legend.key.width = unit(0.5, 'cm')
     )
})


library(cowplot)
combineRocs <- plot_grid(plotlist = singleROCS,
                         labels = "AUTO",
                         label_size = 8,
                         label_x = 0, label_y = 1,
                         nrow=1)

path <- "~/Downloads"
ggsave(plot=combineRocs,
       filename = file.path(path,paste0("threepanelROCs.pdf")),
       units="mm",width=230,height=70)

```

### Grid-based ROC plot

THis section is a work in progress

```{r}
#Load helper functions for ROC curve generation
source("https://raw.githubusercontent.com/ThomasPSpargo/LatentClusterALS/master/04_ML_prediction/roc_func.R")

#Import the xgboost and random forest models in pairs per dataset
ML_grid_pairs<- list(
  ## Clinical all
  clinALL=list(XGB="/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/MLprediction/data_pheno/XGBoost_AUC_maybeFinal/results/XGBfinalTune.Rds",
               RF="/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/MLprediction/data_pheno/RF_AUC_maybeFinal/RFfinalTune.Rds"),
  
  #Genetic models
  genetic=list(XGB="/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/MLprediction/data_rarepathways_allPRS/XGBoost_AUC_maybeFinal/results/XGBfinalTune.Rds",
                RF="/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/MLprediction/data_rarepathways_allPRS/RF_AUC_maybeFinal/RFfinalTune.Rds"),
   
  #Genetic dataset matched clinical models
  clinMatch=list(XGB="/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/MLprediction/phenofilter_rarepathways_allPRS/XGBoost_AUC_maybeFinal/results/XGBfinalTune.Rds",
                 RF="/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/MLprediction/phenofilter_rarepathways_allPRS/RF_AUC_maybeFinal/RFfinalTune.Rds")
)
library(tidyverse)
library(pROC)

#Loop across datasets and generate grid ROC curves
panelplots<- lapply(ML_grid_pairs,function(x){
  
  ## Read in the RF and XGBoost models
  XGBtune<- readRDS(x[[1]]); XGBtune <- XGBtune$pred
  RFtune <- readRDS(x[[2]]); RFtune <- RFtune$pred
  
  #Set a named vector for labelling facets
  labels= setNames(gsub("X","Class ",levels(XGBtune$obs)), levels(XGBtune$obs)) # create named vector
  
  #Supply a list of predictions (one from xgboost and one from RF) for the given dataset
  finalPred=list("xgboost"=XGBtune,"randomForest"=RFtune)
  
  #Run GridROC
  gridCompare<- gridROC(finalPred,labels=labels,plotAUCvals=FALSE)
  
  #Set aesthetics, manually position auc values based on RF and xgboost, and return the plot
  gridAUCplot<- gridCompare$gridROC+
  scale_color_manual(values=c("randomForest"="olivedrab","xgboost"="royalblue"),
                     labels=c("Random forest","eXtreme Gradient Boosting"),name="Algorithm")+
  theme(legend.position = "top")+
    geom_text(data = gridCompare$allAUCs[gridCompare$allAUCs$group=="randomForest",],
              aes(x=0.25,y=0.1,label=signif(AUC,3),colour=group),
              show.legend = FALSE)+
    geom_text(data = gridCompare$allAUCs[gridCompare$allAUCs$group=="xgboost",],
              aes(x=0.25,y=0.3,label=signif(AUC,3),colour=group),
              show.legend = FALSE)
  
  return(gridAUCplot)
  
})
#panelplots


#Save figs
path <- "~/Downloads"
for(i in seq_along(panelplots)){
  ggsave(plot=panelplots[[i]],
          filename = file.path(path,paste0("GridRoc_",names(ML_grid_pairs)[i],".pdf")),
          units="mm",width=160,height=160)
}

```


### SHAP VALUES

Plot absolute mean SHAP values for RF and xgboost algorithms faceted by dataset / algorithm configuration

```{r SHAP}
library(tidyverse)

######
### Chunk setup
######
#Function to get abs. mean shap. Declared as such for mapply circumstance when predicting multiclass probabilities
getMeanSHAP <- function(x,num=1,name){
  x %>%
    as.data.frame() %>%
    summarise(across(everything(),~mean(abs(.)))) %>%
    pivot_longer(cols=everything(),names_to = "Feature") %>%
    mutate(class=num,
           Model=gsub(".*(randomForest|xgb.Booster).*","\\1",name),
           type=if_else(grepl("class12",name),"binary","multiclass"),
           data_config=case_when(grepl("filter",name) ~"Clinical [matched]",
                                 grepl("_pheno",name) ~ "Clinical [full]",
                                 TRUE ~ "Clinical and genetic"
                                 
           )
    ) 
}

######
### RUN
######

shappath<- list.files("/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/MLprediction/weightedShapley",full.names = TRUE)
shapFits<- lapply(shappath,readRDS)
names(shapFits)<-basename(shappath)

#Across each fit, identifying the model by characteristics/names, determine mean shap values
meanShaps <- mapply(function(x,name){

  if("list" %in% class(x$S) && length(x$S)>2){ #For multiclass summaries, mapply across each list element and class number
    
    shapJ <- mapply(getMeanSHAP,x$S,
                    num=paste0("Class ",1:length(x$S)),
                    name=name,
                    SIMPLIFY = FALSE)
    meanShaps <- do.call(rbind,shapJ)
  } else { #For binary summaries apply function once  [note that binary objectives return list class with identical values in opposing directions according to the 'positive' class]
    if("list" %in% class(x$S)){ #
      ShapMatrix <- x$S[[1]]
    } else {
      ShapMatrix <- x$S
    }
    meanShaps <- getMeanSHAP(ShapMatrix,num="Binary",name=name)
  }
  return(meanShaps)
},shapFits,names(shapFits),
SIMPLIFY = FALSE)

plotdata<- do.call(rbind,meanShaps)

#Declare conditions for ordering the features in the resulting figure
refgroup<-plotdata$class=="Binary" &
  plotdata$Model=="xgb.Booster" &
  plotdata$type=="binary" &
  plotdata$data_config == "Clinical and genetic"

### Generate plot data
shapBars_data<- plotdata %>%
  mutate(Feature=recode(Feature,!!!recodes), 
         class = relevel(as.factor(class),ref="Binary"),
         Feature = factor(Feature,levels=Feature[refgroup][order(value[refgroup])])
  )

shapBars <- ggplot(shapBars_data,aes(x = value, y = Feature,fill=Model))+
  facet_grid(cols=vars(class),rows=vars(data_config),scales="free_y",space="free_y")+
  geom_col(position=position_dodge())+
  theme_bw()+
  labs(x = "SHAP value (absolute mean)",y = "Feature")+
  scale_fill_manual(values=c("randomForest"="olivedrab","xgb.Booster"="royalblue"),
                    labels=c("Random\nforest","eXtreme\nGradient\nBoosting"),name="Algorithm")+
  theme(legend.position = "right")
  

#Save fig
path <- "~/Downloads"
ggsave(plot=shapBars,
       filename = file.path(path,paste0("shapImportance.pdf")),
       units="mm",width=250,height=140)

```

### Beeswarm for shapley

This chunk generates beeswarm plots for shap values in order to not simplify only to the average mean. However the plots are quite difficult to condense because of the number of models and are not readily interpretable. Retained in this script for completeness


```{r}
library(ggbeeswarm)

######
### Chunk setup
######

#Function for minmax scaling of features
minmax <- function(x, na.rm = TRUE){
    (x-min(x))/(max(x)-min(x))
}

#Function to prepare the SHAP values
pivotSHAP <- function(piv,num=1,values_to="value",name,scale=FALSE){
  if(scale){piv<- apply(piv,2,minmax)} #If scaling, minmax across columns
  piv %>%
    as.data.frame() %>%  
    pivot_longer(cols=everything(),names_to = "Feature",values_to = values_to) %>%
    mutate(ID=row_number(), #For joining between datasets
           class=num,
           Model=gsub(".*(randomForest|xgb.Booster).*","\\1",name),
           type=if_else(grepl("class12",name),"binary","multiclass"),
           data_config=case_when(grepl("filter",name) ~"Clinical [matched]",
                                 grepl("_pheno",name) ~ "Clinical [full]",
                                 TRUE ~ "Clinical and genetic"
                                 
           )
    ) 
}

######
### RUN
######

#Across each fit, identifying the model by characteristics/names, determine mean shap values
allShaps <- mapply(function(x,name){
  
  #Obtain data to which shapvals correspond - this is always a single matrix
  shapX <-  x$X %>%
    as.data.frame() %>%
    mutate(across(everything(),~minmax(.))) %>%
    pivot_longer(cols=everything(),names_to = "Feature",values_to = "FeatureValue") %>%
    mutate(ID=row_number())
  
  #For multiclass summaries, mapply across each list element and class number
  if("list" %in% class(x$S) && length(x$S)>2){
    
    #Obtain shap values
    shapJ <- mapply(pivotSHAP,x$S,
                    num=paste0("Class ",1:length(x$S)),
                    name=name,
                    values_to="SHAPValue",
                    SIMPLIFY = FALSE) %>%
      lapply(.,full_join,shapX,by=c("Feature", "ID")) %>%
      lapply(.,select,-ID)
    
    SHAPdf <- do.call(rbind,shapJ)
  } else {
    #For binary summaries apply function once [note that binary objectives return list class with identical values in opposing directions according to the 'positive' class]
    if("list" %in% class(x$S)){ matS <- x$S[[1]]
    } else { matS <- x$S }
    
    shapS <- pivotSHAP(matS,num="Binary",name=name,values_to="SHAPValue")
    
    SHAPdf<- full_join(shapS,shapX,by=c("Feature", "ID")) %>%
      select(-ID)
  }
  return(SHAPdf)
},shapFits,names(shapFits),
SIMPLIFY = FALSE)


# ### Generate faceted beeswarm plots for each xgboost algorithm [not run]

# #Concatenate binary and multiclass objectives for Xgboost algorithms
# collect<-list(dataconfig=c("Clinical \\[full\\]","Clinical \\[matched\\]", "Clinical and genetic"),algorithm="xgb.Booster")
# 
# #Concatenate across multi=class and binary models according to data config; keep to only xgboost algorithms
# collectShaps <- list()
# for(i in 1:3){
#   collectShaps[[i]] <- do.call(rbind,allShaps[sapply(allShaps,function(x){
#     any(grepl(collect$dataconfig[i],x$data_config)) & any(grepl(collect$algorithm,x$Model))
#   })])
#   
#   names(collectShaps)[i] <- collect$dataconfig[i]
# }
#
# shapBees<- lapply(collectShaps,function(x){
#   
#   fig<- x %>%
#     #slice_sample(n=1000) %>% #Downsample
#     mutate(Feature=recode(Feature,!!!recodes),
#            Feature=factor(Feature,levels=levels(shapBars_data$Feature)[levels(shapBars_data$Feature) %in% Feature]),
#            class = as.factor(class)
#     ) %>%
#     ggplot(aes(x = SHAPValue, y = Feature,color=FeatureValue))+
#     facet_wrap(~class,nrow=1)+
#     #From ggbeeswarm, quasirandom gives violin-like plots
#     #geom_point(position = position_quasirandom(width=0.4,varwidth = TRUE))+
#     geom_quasirandom(varwidth = TRUE)+
#     geom_vline(xintercept = 0,lty=2)+
#     theme_bw()+
#     labs(x = "SHAP value",y = "Feature",colour="Feature value\n(min-max scaled)")+
#     scale_color_distiller(palette="RdYlBu")+
#     theme(legend.position = "right",
#           plot.title.position = "plot")
#   return(fig)
# })
#
# ## Save to file
#path <- "~/Downloads"
#ggsave(file.path(path,paste0("shapBeeswarm_v4.png")),shapBees$`Clinical and genetic`,height=130,width=210,units="mm")

### Not run syntax to generate a plot with everything all in one figure
# allInOne <- do.call(rbind,collectShaps)
# 
# #Try doing all in one fig
# allInOneFig<- allInOne %>%
#   #slice_sample(n=1000) %>% #Downsample
#   mutate(Feature=recode(Feature,!!!recodes),
#          Feature=factor(Feature,levels=levels(shapBars_data$Feature)[levels(shapBars_data$Feature) %in% Feature]),
#          class = as.factor(class)
#   ) %>%
#     ggplot(aes(x = SHAPValue, y = Feature,color=FeatureValue))+
#     facet_grid(cols=vars(class),rows=vars(data_config),scales="free_y",space="free_y")+
#     #From ggbeeswarm, quasirandom gives violin-like plots
#     #geom_point(position = position_quasirandom(width=0.4,varwidth = TRUE))+
#     geom_quasirandom(varwidth = TRUE)+
#     geom_vline(xintercept = 0,lty=2)+
#     theme_bw()+
#     labs(x = "SHAP value",y = "Feature",colour="Feature value\n(min-max scaled)")+
#     scale_color_distiller(palette="RdYlBu")+
#     theme(legend.position = "right",
#           plot.title.position = "plot")

#ggsave(file.path(path,paste0("shapBeeswarm_allinone.png")),allInOneFig,height=140,width=210,units="mm")

#Generate legends to make clear the contents of each figure
captions <- paste0("Figure SX",seq_along(shapFits),". Feature importance for ",
       ifelse(grepl("xgb.Booster",names(shapFits)),"eXtreme Gradient Boosting","Random Forest"),
       " algorithm predicting class membership ",
       ifelse(grepl("class12",names(shapFits)),"across classes 1 and 2","with a multiclass objective"),
       " in the ",
       ifelse(grepl("rarepathways",names(shapFits)),
              ifelse(grepl("filter",names(shapFits)),"clinical [matched]","clinical and genetic"),
                                                    "clinical [full]"),
       " dataset.",
       ifelse(grepl("class12",names(shapFits))," Positive SHAP values indicate increased predicted probability of being in class 1.","")
)


### Generate beeswarm plot for each ML algorithm with faceting of multiclass variables
shapBees<- mapply(function(x,sub){

  fig<- x %>%
    #slice_sample(n=1000) %>% #Downsample
    mutate(Feature=recode(Feature,!!!recodes),
           class = as.factor(class)
    ) %>%
    ggplot(aes(x = SHAPValue, y = Feature,color=FeatureValue))+
    {if(!grepl("across classes 1 and 2",sub)) facet_wrap(~class,nrow=1)}+
    geom_quasirandom(varwidth = TRUE)+     #From ggbeeswarm, quasirandom gives violin-like plots
    geom_vline(xintercept = 0,lty=2)+
    theme_bw()+
    labs(x = "SHAP value",y = "Feature",colour="Feature value\n(min-max scaled)",subtitle =str_wrap(sub, 140)
           )+
    scale_color_distiller(palette="RdYlBu")+
    theme(legend.position = "right",
          plot.title.position = "plot")
  return(fig)
},allShaps,captions,SIMPLIFY = FALSE)

## Save individual figs to pdf file. Note that this is heavily overplotted and therefore the file is large
path <- "~/Downloads"
pdf(file.path(path,paste0("shapBeeswarm_v5.pdf")),height=6,width=10)
for (i in seq_along(shapBees)) print(shapBees[[i]])
dev.off()


```





##### Descriptive plots

Generate descriptive plots comparing the classes

```{r}
library(survival)
library(survminer)
library(cowplot)

data <- data.table::fread("/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/JNT_JanFinal_savedata_C5/completesample/saved_mplusfit.tsv")

#Prepare list storing continuous variables
ContPlots <- list(DisDur=NULL,AgeOns=NULL,DiagDel=NULL)

#Survival KM
#Palette plot for classes
cbbPalette <- c("gold", "violetred", "#56B4E9", "#000000", "limegreen", "#D55E00", "#CC79A7", "#F0E442")

#Fit survival curve
KMfit<- survfit(Surv(Time_to_death_or_last_followup_years,survival_status_bin) ~ C,data=data)

#Build KM plot
ContPlots[[1]] <- ggsurvplot(KMfit,data=data,
                     pval = F,pval.method=F, conf.int = T, risk.table = F,
                     xlab= "Time (years)",
                     legend="none",
                     palette = cbbPalette,
                     ggtheme = theme_bw()
)[[1]]

#Age of Onset KM
aooKMfit<- survfit(Surv(Age_at_onset_years) ~ C,data=data)
names(aooKMfit$strata)<- gsub("C=","", names(aooKMfit$strata))

#Build KM plot
ContPlots[[2]] <- ggsurvplot(aooKMfit,data=data,
                     pval = F,pval.method=F, conf.int = T, risk.table = F,
                     xlab= "Time (years)",
                     ylab= "1 - Onset probability",
                     legend="top",
                     palette = cbbPalette,
                     ggtheme = theme_bw(),
                     legend.title="Class"
)[[1]]
 


#Diagnostic delay density
ContPlots[[3]]<- data %>%
  filter(!is.na(CTYDEL)) %>%
  mutate(C=as.factor(C)) %>%
ggplot(data=.,aes(CTYDEL,color=C,fill=C))+
  geom_density(alpha=0.7)+
  theme_bw()+
  scale_fill_manual(values=cbbPalette,guide="none")+
  scale_color_manual(values=cbbPalette,guide="none")+
  #guides(fill=guide_legend(override.aes = list(alpha = 1)))+
  labs(x="Diagnostic delay (standardised)",y="Density",fill="Class")


#Diagnostic delay density
CategoryDf<- data %>%
  dplyr::select(C,new_num_id,Sex_at_birth,Phenotype,Site_of_Onset) %>%
  mutate(Sex_at_birth=case_when(Sex_at_birth==1 ~ "Male",
                                Sex_at_birth==2 ~ "Female",
                                TRUE ~ NA_character_),
         Phenotype=case_when(Phenotype==1 ~ "ALS",
                             Phenotype==2 ~ "PLS",
                             Phenotype==3 ~ "PMA",
                             TRUE ~ NA_character_),
         Site_of_Onset=case_when(Site_of_Onset==1 ~ "Other",
                                 Site_of_Onset==2 ~ "Bulbar",
                             TRUE ~ NA_character_),
         C=as.factor(C)) %>%
  reshape2::melt(measure.vars=c("Sex_at_birth","Phenotype","Site_of_Onset")) %>%
  mutate(#C=paste0("Class ", C),
         variable=case_when(variable=="Sex_at_birth" ~ "Sex",
                            variable=="Phenotype" ~ "Clinical\ndiagnosis",
                            variable=="Site_of_Onset" ~ "Site of\nonset"

           ),
         value=as.factor(value)) %>%
  filter(!is.na(value)) #Remove any NAs?

var<- c("Sex","Clinical\ndiagnosis","Site of\nonset")
BarPlots<- lapply(var,function(x,data){
  
  #Create gradient-type scale according to number of classes
  nsteps<- length(unique(data$value[data$variable==x]))
  gradPalette  <- scales::seq_gradient_pal(low = "springgreen", high = "springgreen4", space = "Lab")(1:nsteps/nsteps)
  
  #Generate plot
  data %>%
    filter(variable==x) %>%
    ggplot(data=.,aes(x=C,fill=value))+
    geom_bar(position='fill',alpha=1)+
    labs(fill=x,y="Percentage",x="Class")+
    scale_y_continuous(expand = expansion(mult = c(0,0)), #Set the Y-axis to not truncate
      labels=scales::label_number(scale = 1*100,accuracy = 1))+ #Convert to percentage
    scale_fill_manual(values=gradPalette)+
    theme_bw()+
    #guides(fill=guide_legend(nrow=nsteps-1, byrow=TRUE))+
    theme(legend.position = "bottom",axis.line.y = element_blank())
},CategoryDf)

#Extract and then drop the legend
legend<- get_legend(ContPlots[[2]])
ContPlots[[2]] <- ContPlots[[2]] +
  theme(legend.position = "none")

#Collect the main panels with cowplot
Panels<- plot_grid(plotlist = c(ContPlots,BarPlots),
  labels = "AUTO",
  label_size = 8,
  label_x = 0, label_y = 1
)

#Add the legends
FullDescriptives<- plot_grid(legend,Panels,ncol=1,rel_heights = c(.1, 1))

#Save the figure
path <- "~/Downloads"
ggsave(plot=FullDescriptives,
       filename = file.path(path,paste0("DescriptiveStatFigure.pdf")),
       units="mm",width=230,height=140)
```


### Compare the distribution of people across training-dataset and joint dataset 5-class models

This chunk generates Sankey diagrams across 2-panels. Panel A is the distribution of people across classes from the two models in the training sample (Project MinE). Panel B is the same but for the validation (STRENGTH) data.

```{r}

library(data.table)
library(tidyverse)
library(ggsankey) #https://r-charts.com/flow/sankey-diagram-ggplot2/

#Colour palette for sankey diagram
cbbPalette <- c("gold", "violetred", "#56B4E9", "#000000", "limegreen", "#D55E00", "#CC79A7", "#F0E442")

#Specify paths to the LCA classified datasets
#trainingdata and testdata are based on the Project MinE 5-class model
#jointdata is based on the ProjectMiNE+STRENGTH 5-class model
modelpaths<-c(trainingdata="/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/PM_JanFinal_savedata_C5/completesample/saved_mplusfit.tsv",
  testdata="/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/PM_JanFinal_savedata_C5/completesample/STR_NOPM_ANY_predict_cens_as_is/LCAfit_new_data.tsv",
  jointdata="/Users/tom/OneDrive - King's College London/PhD/PhD project/Latent cluster analysis/Writeup/WriteupModels/JNT_JanFinal_savedata_C5/completesample/saved_mplusfit.tsv")
  
models<- lapply(modelpaths,fread)

#Save data combos in a list
harmoniseC <- list(train=NA,test=NA,unify=NA)

#Extract people from training data and their class assignments from the two models
harmoniseC$train <- left_join(models$trainingdata,models$jointdata,by="new_num_id") %>%
  select("new_num_id","C.x","C.y")

#Extract people from test data and their class assignments from the two models
harmoniseC$test <- left_join(models$testdata,models$jointdata,by="new_num_id") %>%
  select("new_num_id","C.x","C.y")

#Extract people from both training and test data and their class assignments from the two models
harmoniseC$unify <- full_join(models$trainingdata,models$testdata,by=c("new_num_id","C")) %>%
  left_join(models$jointdata,by="new_num_id") %>%
  select("new_num_id","C.x","C.y")

#Tabulate classes in training data and sort largest class to smallest
classOrder<- data.frame(x=names(rev(sort(table(harmoniseC$train$C.x)))),y=NA)
ymatchedIndex <- vector(mode="list",length=nrow(classOrder))

#Loop across subset rows and extract the most likely equal class based on largest overlap. Only search against classes NOT already classified
for(i in 1:nrow(classOrder)){
  xSubgroup<- harmoniseC$train[harmoniseC$train$C.x==classOrder$x[i]]
  tabl <- table(xSubgroup$C.y[!xSubgroup$C.y %in% classOrder$y])
  classOrder$y[i] <- names(tabl)[tabl==max(tabl)]
  ymatchedIndex[[i]] <- harmoniseC$train$C.y==classOrder$y[i]
}

#Prepare a 'total' row of the summary
classOrder[nrow(classOrder)+1,] <- "Total"

#After matching complete, determine class matching properties
#PM indicates the training dataset model
#Joint indicates the 'joint' dataset model
for(j in 1:length(harmoniseC)){  
  propmatchof_Y<- ncol(classOrder)+1
  propmatchof_X<- propmatchof_Y+1
  xClassN<- propmatchof_Y+2
  yClassN<- propmatchof_Y+3
  classOrder[propmatchof_Y:yClassN] <- NA
  names(classOrder)[propmatchof_Y:yClassN] <- paste0(c("Prop_PM_matching_joint_","Prop_Joint_matchingPM","N_PMfit_","N_Jointfit_"),names(harmoniseC)[j],"dataset")
  for(i in 1:nrow(classOrder)){
    
    if(i!=nrow(classOrder)){
      #Calculate proportion of proportion of people assigned to a given class in the joint dataset model with a matching assignment in the PM only model fit
      classOrder[i,propmatchof_Y] <- sum(harmoniseC[[j]]$C.y==classOrder$y[i] & harmoniseC[[j]]$C.x==classOrder$x[i])/sum(harmoniseC[[j]]$C.y==classOrder$y[i])
      classOrder[i,propmatchof_X] <- sum(harmoniseC[[j]]$C.y==classOrder$y[i] & harmoniseC[[j]]$C.x==classOrder$x[i])/sum(harmoniseC[[j]]$C.x==classOrder$x[i])
      classOrder[i,xClassN] <- sum(harmoniseC[[j]]$C.x==classOrder$x[i])
      classOrder[i,yClassN] <- sum(harmoniseC[[j]]$C.y==classOrder$y[i])
    } else {
      classOrder[i,propmatchof_Y] <- sum(paste0(harmoniseC[[j]]$C.y,"_",harmoniseC[[j]]$C.x) %in% paste0(classOrder$y[-i],"_",classOrder$x[-i]))/nrow(harmoniseC[[j]])
      classOrder[i,propmatchof_X] <- classOrder[i,propmatchof_Y]
      classOrder[i,xClassN] <- nrow(harmoniseC[[j]])
      classOrder[i,yClassN] <- nrow(harmoniseC[[j]])
    }
  }
}
names(classOrder)[1:2] <- c("PMfit_class","Jointfit_class")

write.csv(classOrder,file=file.path(dirname(modelpaths[["jointdata"]]),"PMfit_vs_Jointfit.csv"),row.names = FALSE)

#Dataset is the data to plot
#overlapinfo is a 2 column dataframe indicaitng the overlaps between the classes

plotSankey<- function(dataset,overlapinfo){
  #Prepare the Discovery sample for a sankey plot
  sankeyData<- dataset %>%
    mutate(Discovery=as.factor(C.x),
           Joint=as.factor(C.y))
  
  levels(sankeyData$Discovery) <- paste0(levels(sankeyData$Discovery)," [",round(overlapinfo[,1]*100,1),"%]")
  levels(sankeyData$Joint) <- paste0(levels(sankeyData$Joint)," [",round(overlapinfo[,2]*100,1),"%]")
  
  
  sankeyPlot<- sankeyData %>%
    make_long(Discovery,Joint) %>%
    ggplot(., aes(x = x, 
                  next_x = next_x, 
                  node = node, 
                  next_node = next_node,
                  fill = factor(node),
                  label = node)) +
    geom_sankey(flow.alpha = 0.5, node.color = 1,width=0.6) +
    geom_sankey_label(size = 3.5, color = 1, fill = "white") +
    theme_sankey(base_size = 16)+
    scale_fill_manual(values=sapply(cbbPalette,rep,2))+
    scale_x_discrete(position = "top")+
    theme(legend.position = "none",
          axis.title.x = element_blank())
  
  return(sankeyPlot)
}

#Generate Sankey plot for the train data
sankeyTrain <- plotSankey(harmoniseC$train,classOrder[1:5,c("Prop_Joint_matchingPMtraindataset","Prop_PM_matching_joint_traindataset")])

#Generate for the test data
sankeyTest <- plotSankey(harmoniseC$test,classOrder[1:5,c("Prop_Joint_matchingPMtestdataset","Prop_PM_matching_joint_testdataset")])


sankeyFig<- cowplot::plot_grid(sankeyTrain,sankeyTest,nrow=1,labels=c("A","B"))

path <- "~/Downloads"
ggsave(plot=sankeyFig,
       filename = file.path(path,paste0("CompareModelsSankey.pdf")),
       units="mm",width=160,height=160)

```



